Leveraging Diversity to Find Bugs\\ in JavaScript Engines

### 2018-08-02 paper/paper.tex
pass on section 2

JavaScript engines are virtual machines that parse source code and
compile it in bytecodes. These engines implement some version of the
ECMAScript (\es{}), which emerged with the goal to standardize
variants of the language, such as Netscape's JavaScript and
Microsoft's JScript\footnote{The name JavaScript still prevails today,
  certainly for historical reasons.}. The \es{} specification is
regulated by Ecma International~\cite{es6-website} under the
TC39~\cite{tc39-github} technical committee.  Every year, a new
version of the \es{} specification is produced with new features and
minor fixes. The canonical spec today is
ES6~\cite{ecmas262-spec-repo,ecmas262-spec}.

%% In the following, we briefly describe different sources of
%% discrepancies that can emerge across different engine implementations
%% as results of the active changes in the \es{} specification.

.........
A compatibility table relating features and supporting engines is
publicly available~\cite{kangax}.

Another
challenge is undefined (or implementation-dependent) behavior. The
\es{} spec uses the label ``implementation-dependent'' to indicate
cases where each engine developer can decide how to implement some
aspect of a functionality. In these cases, the behavior of some
language element may differ from engine to engine. One reason this
flexibility in the spec exists is to enable compiler
optimizations. For example, the \mcodeid{for-in} loop does not clearly
specify the iteration
order~\cite{so-forin-undefined,javascript-in-chrome} and different
engines capitalize on that~\cite{for-in-undefined}.  As another
example, the latest version of the \es{} specification states that if
the \mcodeid{Number.toPrecision()} function is called with more than
one argument, then the behavior is
implementation-dependent~\cite{es6-toPrecision}. Various other cases
like this exist in the specification.

\subsection{Undefined (Implementation-Dependent) Behavior}

%% \Mar{explain what is an **acceptable** incompatibilities. cite example
%% bug reports.}
%% \Igor{For example, the latest version of ES specs~\cite{es6-spec} describes that
%% the function \CodeIn{Number.toPrecision(precision)}, in general,
%% if precision is defined the method returns a String 
%% containing the Number value represented either 
%% in decimal exponential notation or in decimal 
%% fixed notation. The Number object uses a 
%% property that the specification describes as approximately so when we use 
%% toPrecision with a bigger precision, the values of each engine may vary, this
%% could be considered as a case of incompatibility by engine
%% design.}\Mar{I don't understand the example very well and don't understand the
%%   relationship with HI/LO.}

\subsection{Randomness}

\sloppy Some elements in JS have non-deterministic behavior (\eg{},
Math.random and Date). A test that make decisions based on these
elements could, in principle, produce different outcomes on different
runs. Carefully-written test cases should not manifest this kind of
flakiness~\cite{luo-etal-fse2014,palomba-zaidman-icsme2017}, but
fuzzed test files are more likely to manifest them.

\subsection{\Fix{Other}}

1) The stack depth in JS is often close to the real (native) C stack
depth, nevertheless observable by JS through recursion. This causes
differential behavior even on the same engine when JITs are enabled or
disabled, as the stack frame sizes vary. LangFuzz in particular
produces a lot of recursions and hits these problems even more often
than jsfunfuzz.\Mar{ask C. Holler. it is unclear to me how the output
  could be different.}

\subsection{Bugs}

It is worth to recall that there is no formal specification for \es{};
the specification is written in natural language. As such, differences
in behavior from one engine to another, other than those mentioned
above (Sections~\ref{sec:imp-dep-behavior}
and~\ref{sec:undef-behavior}), can occur. But these discrepancies are
not acceptable.

@@

Figure~\ref{fig:workflow} illustrates the infrastructure we used for
differential testing.  The bug-finding process takes on input JS files
from regression test suites of various JS engines and generates
warnings on output. Boxes in the figure denote encapsulation; arrowed
lines indicate data flow. The cycle icons denote repetition--the
leftmost icon indicates that each file in the input list will be
analyzed in separate whereas the rightmost icon shows that a single
file will be fuzzed multiple times. The bug-finding process works as
follows.

>>>>

JavaScript engines are virtual machines that parse source code,
compile it in bytecodes, and run these bytecodes. These engines
implement some version of the ECMAScript (\es{}), which emerged with
the goal to standardize variants of the language, such as Netscape's
JavaScript and Microsoft's JScript\footnote{The name JavaScript still
  prevails today, certainly for historical reasons.}. The \es{}
specification is regulated by Ecma International~\cite{es6-website}
under the TC39~\cite{tc39-github} technical committee.  Every year, a
new version of the \es{} specification is produced with new features
and minor fixes. The canonical spec today is
ES6~\cite{ecmas262-spec-repo,ecmas262-spec}.

%% In the following, we briefly describe different sources of
%% discrepancies that can emerge across different engine implementations
%% as results of the active changes in the \es{} specification.
%% .........
%% A compatibility table relating features and supporting engines is
%% publicly available~\cite{kangax}.

Both the specification and implementation of JavaScript are
incomplete. Certain parts of the specification are undefined; it is
responsibility of engine engineers to decide how to implement these
functionalities. The JavaScript spec uses the label
``implementation-dependent'' to indicate these cases, where behavior
may differ from engine to engine. One reason this flexibility in the
spec exists is to enable compiler optimizations. For example, the
\mcodeid{for-in} loop does not clearly specify the iteration
order~\cite{so-forin-undefined,javascript-in-chrome} and different
engines capitalize on that~\cite{for-in-undefined}.  As another
example, the latest version of the \es{} specification states that if
the \mcodeid{Number.toPrecision()} function is called with more than
one argument, then the behavior is
implementation-dependent~\cite{es6-toPrecision}. Various other cases
like these exist in the specification. Such gap also exists at the
implementation level. Given the speed the specification changes and
the complexity of the language some features are not fully implemented
as can be observed by the Kangax compatibility table~\cite{kangax}.

\sloppy It is also worth noting that, as in other languages, some
elements in JS have non-deterministic behavior (\eg{},
\mcodeid{Math.random} and \mcodeid{Date}). A test that make decisions
based on these elements could, in principle, produce different
outcomes on different runs. Carefully-written test cases should not
manifest this kind of
flakiness~\cite{luo-etal-fse2014,palomba-zaidman-icsme2017}, but
fuzzed test files could are more likely, intuitively, to manifest
them.

As mentioned previously, all those aspects make testing \js\ engines
more challenging.

%\subsection{Randomness}

%\subsection{\Fix{Other}}

%% 1) The stack depth in JS is often close to the real (native) C stack
%% depth, nevertheless observable by JS through recursion. This causes
%% differential behavior even on the same engine when JITs are enabled or
%% disabled, as the stack frame sizes vary. LangFuzz in particular
%% produces a lot of recursions and hits these problems even more often
%% than jsfunfuzz.\Mar{ask C. Holler. it is unclear to me how the output
%%   could be different.}

@@

Figure~\ref{fig:workflow} illustrates the infrastructure we used for
differential testing, which is related to part of the study discussed
in this paper.  The bug-finding process takes on input JS files from
regression test suites of various JS engines and generates warnings on
output. Boxes in the figure denote encapsulation; arrowed lines
indicate data flow. The cycle icons denote repetition--the leftmost
icon indicates that each file in the input list will be analyzed in
separate whereas the rightmost icon shows that a single file will be
fuzzed multiple times. The bug-finding process works as follows.
